---
title: "how-to-use"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{how-to-use}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(RepliSeq)
```

# readRS(paths_data,fractions) :

This functions reads Repli-seq assays from multiple files (one file for one fraction) and outputs a dataframe from it.
It requires bedgraph inputs with a one line header listing options, with **no other comments** such as : [(bedgraph spec)](http://genome.ucsc.edu/goldenPath/help/bedgraph.html)   
track 	type=bedGraph 	name=NT_chr22-s1	description=50kbprofile   
chr22	0	50000	0   
chr22	50000	100000	0   

```{r}

### args :
temp_paths <- c("../inst/extdata/NT_chr22-s1.bdg","../inst/extdata/NT_chr22-s2.bdg",
                "../inst/extdata/NT_chr22-s3.bdg","../inst/extdata/NT_chr22-s4.bdg",
                "../inst/extdata/NT_chr22-s5.bdg","../inst/extdata/NT_chr22-s6.bdg")
temp_fractions <- c("S1","S2","S3","S4","S5","S6")

### 2 fractions RepliSeq
# apply function :
RS_early <- readRS(temp_paths[1:2],temp_fractions[1:2])
# outputs :
tail(RS_early)
typeof(RS_early)
summary(RS_early)

### 6 fractions RepliSeq
# apply function :
RS_all <- readRS(temp_paths,temp_fractions)
# outputs :
tail(RS_all)

### 1 fraction RepliSeq ( for S0 controls )
# apply function : 
RS_S0 <- readRS("../inst/extdata/NT_chr22-s0.bdg","S0")
# outputs :
tail(RS_S0)

```

# rescaleRS(rs_assay,scale_factor) :

This functions rescales Repli-seq assays loaded with readRS to lower resolution ( e.g. from 1kb windows to 50kb windows or 100kb)

```{r}

# To rescale from 50kb to 100kb the scale_factor is 2 ( from 1kb to 50kb it would be 50)

### 2 fractions RepliSeq
# apply function :
RS_early_100kb <- rescaleRS(RS_early,2)
# outputs :
tail(RS_early_100kb)

### 6 fractions RepliSeq
# apply function :
RS_all_100kb <- rescaleRS(RS_all,2)
# outputs :
tail(RS_all_100kb)

### 1 fraction RepliSeq ( for S0 controls )
# apply function :
RS_S0_100kb <- rescaleRS(RS_S0,2)
# outputs :
tail(RS_S0_100kb)

```

# smoothRS(rs_assay,scale_factor) :

This functions rescales Repli-seq assays loaded with readRS to lower resolution ( e.g. from 1kb windows to 50kb windows or 100kb)

```{r}

# To get 50kb smoothed profiles from 1kb adjacent profiles smooth_factor = 50 --> results in count matrices with 
# genomic windows having a step = 1kb and a span of 50kb.
# From 50kb to 100kb the smooth_factor is 2 

### 2 fractions RepliSeq
# apply function :
RS_early_smoothed_2 <- smoothRS(RS_early,2)
# outputs :
tail(RS_early_smoothed_2)

### 6 fractions RepliSeq
# apply function :
RS_all_100kb <- rescaleRS(RS_all,2)
# outputs :
tail(RS_all_100kb)

### 1 fraction RepliSeq ( for S0 controls )
# apply function :
RS_S0_100kb <- rescaleRS(RS_S0,2)
# outputs :
tail(RS_S0_100kb)

```

# weight_phases(rs_assay, G1_peak,G2_peak,gates,SimSPhase=NA,TotalReads=NA) :

```{r}

### 6 fractions RepliSeq
# calculate a normalization factor for each phase based on the experimental FACS profile and a simulation of S-phase progressio (Percentage of replicated DNA, Number of active forks):
norm_ratios <- weight_phases(rs_assay = RS_all,
                             G1_peak = 24.15,
                             G2_peak = 44.2,
                             gates = data.frame( start=c(24.2519169380677,27.282955036018755,30.61406625807228,34.19265411124849,37.47504529876629,40.60739992423285),
    end=c(27.282955036018755,30.61406625807228,34.19265411124849,37.47504529876629,40.60739992423285,47.27059148295092)),
    TotalReads = 2000*6 )

print(norm_ratios)

```

# normalizeRS(rs_assay,norm_ratios) :

```{r}
### 6 fractions RepliSeq
# As alternative to the use of weight_phases
# 1st calculate ratios for a sum == 2000 :
norm_ratios <- c(sum(RS_all$S1) / 2000,
                 sum(RS_all$S2) / 2000,
                 sum(RS_all$S3) / 2000,
                 sum(RS_all$S4) / 2000,
                 sum(RS_all$S5) / 2000,
                 sum(RS_all$S6) / 2000)

# apply normalizeRS:
RS_all_normlized<- normalizeRS(RS_all,norm_ratios)
# outputs :
tail(RS_all_normlized)

```



# calculateS50(rs_assay) :

```{r}

temp_rs <- data.frame(chr = rep("chr1",7),
                      start = seq(0,6000,1000),
                      stop = seq(1000,7000,1000),
                      S1 = c(0,0,0,1,1,1,1),
                      S2 = c(0,0,1,1,1,1,0),
                      S3 = c(0,1,1,1,1,0,0),
                      S4 = c(1,1,1,1,0,0,0))


temp_S50 <- RepliSeq::calculateS50(temp_rs)
print(temp_S50)

### 2 fractions RepliSeq
# apply function :
temp.S50 <- RepliSeq::calculateS50(RS_early)
# outputs :
tail(temp.S50)

### 6 fractions RepliSeq
# apply function :
temp.S50 <- RepliSeq::calculateS50(RS_all)
# outputs :
tail(temp.S50)
summary(temp.S50$S50)

```


# calculateURI(rs_x, rs_y) :

This functions calculates URI between two Repli-seq assays. It returns a dataframe with the following columns :   
chr,start,stop,sum_x,sum_y,mean_xy,URI


```{r}
####### load second Repli-seq assay for comparison :
### args :
aph_paths <- c("../inst/extdata/Aph_chr22-s1.bdg","../inst/extdata/Aph_chr22-s2.bdg",
               "../inst/extdata/Aph_chr22-s3.bdg","../inst/extdata/Aph_chr22-s4.bdg",
               "../inst/extdata/Aph_chr22-s5.bdg","../inst/extdata/Aph_chr22-s6.bdg")
aph_fractions <- c("S1","S2","S3","S4","S5","S6")

# read :
RS_aph_all <- readRS(aph_paths,aph_fractions)
RS_aph_early <- readRS(aph_paths[1:2],aph_fractions[1:2])
#######

### 2 fractions RepliSeq
# apply function :
aph_nt_uri <- calculateURI(RS_aph_early,RS_early)
# outputs :
tail(aph_nt_uri)

### 6 fractions RepliSeq
# apply function :
aph_nt_uri <- calculateURI(RS_aph_all,RS_all)
# outputs :
tail(aph_nt_uri)



```


# calculateNoiseRatios(rs_assay, rs_control) :

```{r}

### 2 fractions RepliSeq
# apply function :
ratios_early <- calculateNoiseRatios(RS_early, RS_S0)
# outputs :
print(ratios_early)

### 6 fractions RepliSeq
# apply function :
ratios_all <- calculateNoiseRatios(RS_all, RS_S0)
# outputs :
print(ratios_all)

```

# removeNoise(rs_assay, rs_control, noise_ratios) :

```{r}

### 2 fractions RepliSeq
# apply function :
RS_noise_rm_early <- removeNoise(RS_early, RS_S0,ratios_early)
# outputs :
tail(RS_noise_rm_early)

### 6 fractions RepliSeq
# apply function :
RS_noise_rm_all <- removeNoise(RS_all, RS_S0,ratios_all)
# outputs :
tail(RS_noise_rm_all)

```

# doubleXchr(rs_assay) :

This function can be used to double the values in count matrices at chrX as for example when normalization haven't been performed on raw data from human male samples.

```{r}

### 2 fractions RepliSeq
# apply function :
RS_early_doubledX <- doubleXchr(RS_early,chr = "chr22") # default chr parameter is "chrX"
# outputs :
tail(RS_early_doubledX)

### 6 fractions RepliSeq
# apply function :
RS_all_doubledX <- doubleXchr(RS_all,chr = "chr22") # default chr parameter is "chrX"
# outputs :
tail(RS_all_doubledX)

### 1 fraction RepliSeq (for S0 controls)
# apply function :
RS_S0_doubledX <- doubleXchr(RS_S0,chr = "chr22") # default chr parameter is "chrX"
# outputs :
tail(RS_S0_doubledX)

```

# writeBedgraph(rs_assay,path_file,sample_name) :

```{r}

### 2 fractions RepliSeq
writeBedgraph(RS_early,"../inst/extdata/output","NT_early")

### 6 fractions RepliSeq
writeBedgraph(RS_all,"../inst/extdata/output","NT_all")

### 1 fraction RepliSeq (for S0 controls)
writeBedgraph(RS_S0,"../inst/extdata/output","NT")

```


# writeBigwig(rs_assay,path_file,sample_name,chromsizes,wiggle_start,wiggle_step_wiggle_span) : 

**(requires [wigToBigWig](https://www.encodeproject.org/software/wigtobigwig/) installed on the system)**.   
This function writes Bigwig profiles for all the fractions of a provided Repli-seq assay. It requires UCSC's WigToBigWig installed to convert from Wiggle to BigWig.

```{r}

### 2 fractions RepliSeq
writeBigwig(RS_early,"../inst/extdata/output","NT_early","../inst/extdata/Hg19/chromsizes.txt",wiggle_start = 1,wiggle_step = 50000, wiggle_span = 50000)

### 6 fractions RepliSeq
writeBigwig(RS_all,"../inst/extdata/output","NT_all","../inst/extdata/Hg19/chromsizes.txt",wiggle_start = 1,wiggle_step = 50000, wiggle_span = 50000)

### 1 fraction RepliSeq (for S0 control)
writeBigwig(RS_S0,"../inst/extdata/output","NT","../inst/extdata/Hg19/chromsizes.txt",wiggle_start = 1,wiggle_step = 50000, wiggle_span = 50000)

```


